Collocated Deduplication with Fault Isolation for Virtual Machine Snapshot Backup

### 2013-06-20 snapshot_system/snapshot.tex
Added CDS coverage graph with measured vs predicted. Not sure the predicted is helpful to the graph though.

A deduplication scheme compares the fingerprints of the current snapshot
with its parent snapshot and also other snapshots in the entire cloud.
The tradition approach compares all fingerprints and 
reserve one copy for all of its deduplicates.
We call this as  the VM-oblivious (VO) approach.
In comparison, we analyze the design of
a VM-centric approach (VC) which differenates duplicates within a VM and cross VMs,
and conducts \textit{Inner-VM} and \textit{cross-VM} detection seperately.

@@

Inner-VM duplication can be effecive within  VM's snapshots there
are a large amount of duplicates existing among the snapshots within
each VM.  Localizing the snapshot data deduplication witin  a VM first
increases data independency between different VM backups,
simplifies snapshot management and statistics collection,
and facilitates parallel execution of snapshot operations.
On the other hand, cross-VM duplication can also be desirable mainly due to widely-used software and libraries. 
As the result, different VMs tend to backup large amount of highly similar data.
We integrate those strategies together as follows.
%Our multi-level pipeline process can minimize 
%the cost of deduplication while maximize the its efficiency at each level,
%and it is parallel since each segment is processed independently.

@@

It should be noted that changed block tracking is supported or can be easily implemented in several 
major virtualization solution vendors. 
%(VMware, Xen, Microsoft).  
VMware support it directly. Their hypervisor has API to let external backup application know 
the changed areas since last backup.  Xen doesn't directly support it. However, their open-source 
architecture allows anyone to extend the device driver, thus enabling changed block tracking. 
We implements this way in the Alibaba's platform.  
Microsoft SDK provides an API that allows external application monitor 
the VM's I/O traffic.  Therefore changed block tracking can be implemented externally.

\item \textbf{Chunk-level fine-grain inner-VM detection.}
The best deduplication uses a nonuniform chunk size 
in the average of 4K or 8K~\cite{??}.
Thus the second-level inner-VM deduplication is to assess in this
level, but only for those dirty  segments. 
We load the fingerprints of block chunks of the corresponding segment from the
parent for a comparisons and compare near-by fingerprint matching within this segement.
The amount of memory for maintaining those fingerprints  is small.
For example, with a 2MB segment, there are about 500 fingerprints to compare.

@@

\item \textbf{Cross-VM deduplication.}
This step accomplishes the standand global fingerprint  comparison as conducted
in the previous work~\cite{??}.
One key observation is that the inner deduplication has removed many of duplicates.
There are not lot of deduplication opportunities cross VMs while the memory
consumption for global comparison is expensive.
Thus our approximation is that the global fingerprint  comparison  only searches for the most popular items. 
%Since duplicate sharing patterns among  VM follows a zip-like distribution,  
\end{itemize}

When a block is not detected as a duplicate to any existing block, this block will be written
to the file system. Since the backend file system typically uses a large block size such as 64MB, each VM will 
accumnulate small local blocks. We manage  this accmulation process using an apendstore  scheme
and discuss this in details in Section~\ref{sect:arch}.

 \subsection{Popular Chunk Management}
%VM-oriented Fault Isolation}
%Our objective for fault isolation is to minimize the number of VMs affected when there are failures
%in the cluster.  
%The inner-VM deduplication does not depend on any global service and the comparison
%for each VM is localized within the parent and the current snapshot.
%Lack of data dependence between VMs brings 
With cross-VM deduplication, shared data blocks
create an artifical dependence  among VMs and failure of one shared block
affects many VMs. Thus we only maintain the index for a small set of popular chunks.
The mnagement for popular data chunks contains two aspects.
\begin{itemize}
\item Compute and select top-$k$ most popular chunks.
The popularity of a chunk is the number  of data blocks from different VMs
that are duplicates of this block after the inner VM deduplication.
This number can be computed perioidically in a weekly basis.
Once the popularity of all data blocks is collected, the system only maintains the top $k$
most popular blocks.  For cross-VM comparison,
we only store  top $k$ items and $k$ is  chosen to be small while 
our analysis belows will show that the algorithm can still deliver competitive deduplication effiency.
\item  Since $k$ is small and these top $k$ blocks are shared among multiple VMs, 
we can afford to provide extra replicas for these popular blocks to enhanace the fault resilence.
We will provide an alaysis in the next subsection.
\end{itemize}

This section analyzes how value   $k$ impacts on the deduplication efficiency.
%It also  compares the  fault resilience of our VM-centric deduplication approach with a standand approach using 
%global deduplication.
%\subsubsection{Impact of CDS deduplication}
The analysis is based on the our study 
using Alibaba's production user data~\cite{ieeecloud} 
which shows that the popularity of data blocks after inner VM deduplication follows 
a Zipf-like distribution\cite{zipf} and its
exponent $\alpha$ is ranged between 0.65  and  0.70.
%As a result, it can be proved that deduplication efficiency of CDS index is scalable:

@@

Let $b$ be the total amount of data blocks after localized deduplication, 
$b_u$ be the total amount of fingerprints 
in the global index after complete deduplication, 
$C_i$ be the frequency for the $i$th most popular fingerprint. 
By Zipf-like distribution, $C_i = \frac{C_1}{i^\alpha}.$
Since $ \sum_{i=1}^{b_u}C_i = b$,
\[
C_1 \sum_{i=1}^{b_u}\frac{1}{i^\alpha} = b
\]
Given $\alpha <1$, $C_1$ can be approximated with integration:
\begin{equation}
C_1=\frac{b(1-\alpha)}{b_u^{1-\alpha}}.
\end{equation}

@@

>>>>

A deduplication scheme compares the fingerprints of the current snapshot
with its parent snapshot and also other snapshots in the entire cloud.
The traditional approach compares all fingerprints and 
stores one copy for all of a block's duplicates.
We call this as  the VM-oblivious (VO) approach.
In comparison, we analyze the design of
a VM-centric approach (VC) which differenates duplicates within a VM and cross VMs,
and conducts \textit{Inner-VM} and \textit{cross-VM} detection seperately.

@@

Inner-VM duplication can be very effecive within VM's snapshots. There
are typically a large number of duplicates existing among the snapshots
for a single VM.  Localizing the snapshot data deduplication witin a VM
improves the system by:
increasing data independency between different VM backups,
simplifying snapshot management and statistics collection,
and facilitating parallel execution of snapshot operations.
On the other hand, cross-VM duplication can also be desirable mainly due to widely-used software and libraries. 
As the result, different VMs tend to backup large amount of highly similar data.
We integrate those strategies together as follows.
%Our multi-level pipeline process can minimize 
%the cost of deduplication while maximize the its efficiency at each level,
%and it is parallel since each segment is processed independently.

@@

It should be noted that changed block tracking is supported or can be easily implemented in many 
major virtualization solution vendors. 
%(VMware, Xen, Microsoft).  
VMware support it directly; the VMWare hypervisor has an API to let external backup application know 
the changed areas since last backup. Xen doesn't directly support it, However, their open-source 
architecture allows anyone to extend the device driver, thus enabling changed block tracking. 
We implement dirty bit tracking this way in Alibaba's platform.  
The Microsoft SDK provides an API that allows external applications to monitor 
the VM's I/O traffic, therefore changed block tracking can be implemented externally.

\item \textbf{Chunk-level fine-grain inner-VM detection.}
The best deduplication uses a nonuniform chunk size 
in the average of 4K or 8K~\cite{??}. This allows the system to
achieve deduplication even when there are insertions/deletions which would
affect many fixed-size blocks.
Thus the second-level inner-VM deduplication is to perform this chunking for
dirty segments, and to compare the snapshot with its parent. 
We load the fingerprints of block chunks of the corresponding segment from the
parent and perform fingerprint matching for further inner-VM deduplication.
The amount of memory for maintaining those fingerprints  is small, as we only
load one segment at a time.
For example, with a 2MB segment, there are about 500 fingerprints to compare.

@@

\item \textbf{Cross-VM deduplication.}
This step accomplishes the standand global fingerprint  comparison as conducted
in the previous work~\cite{??}.
One key observation is that the inner deduplication has removed many of the duplicates.
There are fewer deduplication opportunities across VMs while the memory and network
consumption for global comparison is more expensive.
Thus our approximation is that the global fingerprint comparison only searches for the most popular items. 
%Since duplicate sharing patterns among  VM follows a zip-like distribution,  
\end{itemize}

When a block is not detected as a duplicate to any existing block, this block will be written
to the file system. Since the backend file system typically uses a large block size such as 64MB, each VM will 
accululate small local blocks. We manage this accumulation process using an apendstore  scheme
and discuss this in details in Section~\ref{sect:arch}.

 \subsection{Popular Chunk Management}
%VM-oriented Fault Isolation}
%Our objective for fault isolation is to minimize the number of VMs affected when there are failures
%in the cluster.  
%The inner-VM deduplication does not depend on any global service and the comparison
%for each VM is localized within the parent and the current snapshot.
%Lack of data dependence between VMs brings 
With cross-VM deduplication, shared data blocks
create an artifical dependence  among VMs and failure of one shared block
affects many VMs. Thus we only maintain the index for a small set of popular chunks.
The management for popular data chunks contains two aspects.
\begin{itemize}
\item Compute and select top-$k$ most popular chunks.
The popularity of a chunk is the number  of data blocks from different VMs
that are duplicates of this block after the inner VM deduplication.
This number can be computed perioidically in a weekly basis.
Once the popularity of all data blocks is collected, the system only maintains the top $k$
most popular blocks.  For cross-VM comparison,
we only store  top $k$ items and $k$ is chosen to be relatively small.
Our analysis below will show that the algorithm can still deliver competitive
deduplication effiency after making these approximations.
\item  Since $k$ is small and these top $k$ blocks are shared among multiple VMs, 
we can afford to provide extra replicas for these popular blocks to enhanace the fault resilence.
We will provide an analysis of the fault tolerance in the next subsection.
\end{itemize}

This section analyzes how value   $k$ impacts the deduplication efficiency.
%It also  compares the  fault resilience of our VM-centric deduplication approach with a standand approach using 
%global deduplication.
%\subsubsection{Impact of CDS deduplication}
The analysis is based on VM traces collected
using Alibaba's production user data~\cite{ieeecloud} 
which shows that the popularity of data blocks after inner VM deduplication follows 
a Zipf-like distribution\cite{zipf} and its
exponent $\alpha$ is ranged between 0.65  and  0.70.
%As a result, it can be proved that deduplication efficiency of CDS index is scalable:

@@

Let $b$ be the total number of data blocks after localized deduplication, 
$b_u$ be the total number of fingerprints 
in the global index after complete deduplication, and
$C_i$ be the frequency for the $i$th most popular fingerprint. 
By Zipf-like distribution, $C_i = \frac{C_1}{i^\alpha}.$
Since $ \sum_{i=1}^{b_u}C_i = b$,
\[
C_1 \sum_{i=1}^{b_u}\frac{1}{i^\alpha} = b
\]
Given $\alpha <1$, $C_1$ can be approximated with integration:
\begin{equation}
C_1=\frac{b(1-\alpha)}{b_u^{1-\alpha}}.
\end{equation}

@@

\begin{figure}[ht]
  \centering
    \begin{tikzpicture}
            \begin{axis}[
            %title={CDS Coverage},
            xlabel={Total Data Stored (GB)},
            ylabel={CDS Coverage (\%)},
            %extra y ticks={4.5,5.5,6.5} %to add extra ticks
            %legend pos=outer north east,
            legend columns=2,
            legend style={
                at={(0.5,-0.2)},
            anchor=north}
            ]
            \addplot[color=blue,mark=*] table[x=InputSize,y=A1] {figures/cds_coverage.txt};
            \addplot[color=blue,mark=square*] table[x=InputSize,y=A2] {figures/cds_coverage.txt};
            \addplot[color=blue,mark=triangle*] table[x=InputSize,y=A4] {figures/cds_coverage.txt};
            \addplot[densely dashed,color=red,mark=*] table[x=InputSize,y=T1] {figures/cds_coverage.txt};
            \addplot[densely dashed,color=red,mark=square*] table[x=InputSize,y=T2] {figures/cds_coverage.txt};
            \addplot[densely dashed,color=red,mark=triangle*] table[x=InputSize,y=T4] {figures/cds_coverage.txt};
            \legend{Measured (1\%CDS),Measured (2\%CDS),Measured (4\%CDS),Predicted (1\%CDS),Predicted (2\%CDS),Predicted (4\%CDS)};
            \end{axis}
    \end{tikzpicture}
  \caption{fixed-alpha predicted vs. actual CDS coverage as data size increases.}
  \label{fig-vo-links}
\end{figure}
